---
title: "Alicia"
author: "Julin Maloof"
date: "2023-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

And here is code for generating a test dataset in the format of my real dataset (the merging of separate data frames is to mimic the process I need to follow for my data):

```{r}
X1 <- data.frame('pos.id' = c("chr1a1", "chr1a2", "chr2a1", "chr2a2", "chr3a1", "chr1b2", "chr3c2", "chr5d1"), 'X1' = c(1, 0, .5, 1, 0, .5, 1, 1))
X2 <- data.frame('pos.id' = c("chr1a1", "chr2a2", "chr3a1", "chr4a2", "chr5a1", "chr5d1", "chr6c2", "chr6d1"), 'X2' = c(1, 0, .5, 1, 0, .5, .5, 1))
X3 <- data.frame('pos.id' = c("chr1a1", "chr1a8", "chr2a7", "chr2a6", "chr3a4", "chr1b3", "chr6c2", "chr5d1"), 'X3' = c(1, 0, .5, 1, 0, .5, 1, 1))
X4 <- data.frame('pos.id' = c("chr1a1", "chr1a2", "chr3a1", "chr4a2", "chr3a4", "chr5d1", "chr6c2", "chr6d2"), 'X4' = c(1, 0, .5, 1, 0, .5, 1, 0))
X5 <- data.frame('pos.id' = c("chr1a1", "chr1a2", "chr2a2", "chr2a3", "chr3a1", "chr5b3", "chr6c2", "chr6d2"), 'X5' = c(1, 0, .5, 1, 0, .5, 1, 1))
X6 <- data.frame('pos.id' = c("chr1a1", "chr1a2", "chr2a1", "chr3a2", "chr3a8", "chr4b2", "chr5c2", "chr6d2"), 'X6' = c(1, 0, .5, 1, 0, .5, 1, 0))
dflist <- list(X1, X2, X3, X4, X5, X6)
allcpg <- dflist %>% reduce(full_join, by='pos.id')
#rm(dflist)
allcpg <- t(allcpg)
head(allcpg[,1:10])
```


Here is my code for filtering my data. The idea is to first filter columns based on number of NAs and then create a list of row names based on number of NAs:
```{r}
allcpg2 <- data.frame(row.names = c("pos.id","X1", "X2", "X3", "X4", "X5", "X6"))
j <- 0
for (i in 1:length(allcpg[1,])){
  if (sum(is.na(allcpg[,i])) < 3){
    allcpg2 <- cbind(allcpg2, allcpg[,i])
    j <- j + 1
  }
}
colnames(allcpg2) <- as.character(allcpg2[1,])
allcpg2 <- allcpg2[-1,]
j
```


```{r}
# identify max number of sites in a sample, aka max number of non-NA entries in a row
namax <- max(apply(allcpg2, MARGIN = 1, FUN = function(x) sum(is.na(x))))
namin <- min(apply(allcpg2, MARGIN = 1, FUN = function(x) sum(is.na(x))))
maxsites <- length(allcpg2[1,]) - namin
# calculate 3% of the max
threshold <- maxsites*.5 # temporarily changed the percentage for the test dataset
# identify samples that have less than 3% max non-NA entries
filtersamples <- function(x){
  sites <- length(allcpg2[1,]) - sum(is.na(x))
  if (sites < threshold){
    x
  }
}
sitelist <- apply(allcpg2, MARGIN = 1, FUN = filtersamples)
sitelist <- sitelist %>% discard(is.null)
sitelist <- names(sitelist)
```

## JM tidy
```{r}
library(tidyverse)
```


Create allcpg but keep it numeric and with column names
```{r}
allcpg_jm <- dflist %>% reduce(full_join, by='pos.id') %>%
  column_to_rownames(var="pos.id") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var="sample_ID")
  
allcpg_jm[,1:5]
allcpg[,1:5]
```

Filter to keep columns with less than 3 NAs
```{r}
allcpg2_jm <- allcpg_jm %>%
  select(where(~ sum(is.na(.x)) < 3))

allcpg2_jm
allcpg2
```
Calculate maxsites and then threshold
```{r}
sitelist_jm <- allcpg2_jm %>%
  rowwise() %>%
  mutate(n.sites=sum(!is.na(c_across(starts_with("chr"))))) %>%
  ungroup() %>%
  filter(n.sites < max(n.sites)*.5) %>%
  pull(sample_ID)

sitelist_jm
sitelist
```

