---
title: "Chapter 14"
author: "Julin Maloof"
date: "2023-04-07"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R6

__"in most cases R6 will lead you to non-idiomatic R code."__

```{r}
library(R6)
```

Use the `R6Class()` function to define a class.  Assign it to an object that has the same name as the class.  First argument is class name, second argument is public fields and methods.

```{r}
Accumulator <- R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)
```

```{r}
Accumulator
```

Create a new object of this clas with $new function

```{r}
x <- Accumulator$new()
```

```{r}
x$sum
x$add(4)
x$sum
```

## 14.2.1 Method Chaining
 
side-effect R6 methods (those that update a field and are not called to create an output) should return self invisibly

then you can do things like this
```{r}
x$
  add(10)$
  add(10)$
  sum
```

## 14.2.2 Important methods

Generally should define $initialize and $print methods to override the defaults

```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) {
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    invisible(self)
  }
))

hadley2 <- Person$new("Hadley")
hadley2
```

## 14.2.3 Adding methods after creation

Use $set

```{r}
Accumulator <- R6Class("Accumulator")
Accumulator$set("public", "sum", 0)
Accumulator$set("public", "add", function(x = 1) {
  self$sum <- self$sum + x 
  invisible(self)
})
```

Will only impact newly created objects, will not be added to previosly created objects of the same class.

## 14.2.5 Inheritance

```{r}
AccumulatorChatty <- R6Class("AccumulatorChatty", 
  inherit = Accumulator,
  public = list(
    add = function(x = 1) {
      cat("Adding ", x, "\n", sep = "")
      super$add(x = x)
    }
  )
)

x2 <- AccumulatorChatty$new()
x2$add(10)$add(1)$sum
```

## 14.2.5 Introspections

Every R6 object has an S3 class hierarchy
```{r}
class(hadley2)
```

## 14.2.6. Exercises

### 1 Create a bank account R6 class that stores a balance and allows you to deposit and withdraw money. Create a subclass that throws an error if you attempt to go into overdraft. Create another subclass that allows you to go into overdraft, but charges you a fee.

```{r}
BankAccountBasic <- R6Class("BankAcountSimple", list(
  balance = 0,
  deposit = function(amt) {
    self$balance <- self$balance + amt
    invisible(self)
  },
  withdraw = function(amt) {
    self$balance <- self$balance - amt
    invisible(self)
  },
  print = function() {
    cat("The current balance is: ", self$balance, "\n")
    invisible(self)
  }
))

BA1 <- BankAccountBasic$new()

BA1$deposit(100)
BA1$withdraw(33)
BA1
```

```{r, error=TRUE}
BankAccount2 <- R6Class("BankAcount2",
                       inherit = BankAccountBasic,
                       public = list(
                         withdraw=function(amt) {
                           stopifnot(self$balance-amt>=0)
                           super$withdraw(amt)
                         }
                       ))

BA2 <- BankAccount2$new()

BA2$deposit(100)
BA2$withdraw(33)
BA2
BA2$withdraw(100)
BA2
```
```{r, error=TRUE}
BankAccount3 <- R6Class("BankAcount3",
                       inherit = BankAccountBasic,
                       public = list(
                         withdraw=function(amt) {
                           super$withdraw(amt)
                           if (self$balance < 0) {
                             warning("New balance is less than $0, a $5.00 fee is being assessed\n")
                             super$withdraw(5)
                           }
                         }
                       ))

BA3 <- BankAccount3$new()

BA3$deposit(100)
BA3$withdraw(150)
BA3
```

### 2. Create an R6 class that represents a shuffled deck of cards. You should be able to draw cards from the deck with $draw(n), and return all cards to the deck and reshuffle with $reshuffle(). Use the following code to make a vector of cards.

```{r}
deck <- R6Class("deck", public = list(
  stack=character(),
  drawn=character(),
  initialize = function() {
    suit <- c("♠", "♥", "♦", "♣")
    value <- c("A", 2:10, "J", "Q", "K")
    self$stack <- sample(paste0(rep(value, 4), suit))
    self$drawn <- character()
  },
  reshuffle = function() {
    self$stack <- c(self$stack, self$drawn)
    self$drawn <- character()
    self$stack <- sample(self$stack)
  },
  draw = function(n) {
    newdraw <- self$stack[1:n]
    self$drawn <- c(self$drawn, newdraw)
    self$stack <- self$stack[-1:-n]
    cat(newdraw)
  },
  print = function() {
    cat("Not Drawn: ",self$stack, "\n")
    cat("Drawn: ", self$drawn, "\n")
  }
))

mydeck <- deck$new()
mydeck

mydeck$draw(5)
mydeck

mydeck$draw(5)
mydeck

mydeck$reshuffle()
mydeck

```
### 3. Why can’t you model a bank account or a deck of cards with an S3 class?

Beacuse S3 is not modify in place

### 4. Create an R6 class that allows you to get and set the current time zone. You can access the current time zone with Sys.timezone() and set it with Sys.setenv(TZ = "newtimezone"). When setting the time zone, make sure the new time zone is in the list provided by OlsonNames().

```{r, error=TRUE}
TZ <- R6Class("TZ", public=list(
  tz=Sys.timezone(),
  get=function() self$tz,
  set=function(new_tz) {
    stopifnot(new_tz %in% OlsonNames())
    Sys.setenv(TZ = new_tz)
    self$tz <- Sys.timezone()
  }
))

tz <- TZ$new()
tz$get()
tz$set("US/Hawaii") # I wish
tz$get()
tz$set("Mars")
tz$get()
```

### 5. Create an R6 class that manages the current working directory. It should have $get() and $set() methods.

```{r, error=TRUE}
WD <- R6Class("WD", public=list(
  wd=getwd(),
  get=function() self$wd,
  set=function(new_wd) {
    stopifnot(dir.exists(new_wd))
    setwd(new_wd)
    self$wd <- new_wd
  }
))

wd <- WD$new()
wd$get()
wd$set("../") 
wd$get()
wd$set("Mars")
wd$get()
wd$set("/Users/jmaloof/git/AdvancedR")
wd$get()
```

### 5. Why can’t you model the time zone or current working directory with an S3 class?

no modify in place?

### 6. What base type are R6 objects built on top of? What attributes do they have?

```{r}

```

## 14.3 Controlling Access

R6class() has two other arguments:

*`private` create fields and methods that are only available inside the class
* `active` to define dynamic or active fields (whatever that means)

### 14.3.1 Privacy

* use the private argument just like public
* access using private$ instead of self$

```{r}
Person <- R6Class("Person", 
  public = list(
    initialize = function(name, age = NA) {
      private$name <- name
      private$age <- age
    },
    print = function(...) {
      cat("Person: \n")
      cat("  Name: ", private$name, "\n", sep = "")
      cat("  Age:  ", private$age, "\n", sep = "")
    }
  ),
  private = list(
    age = NA,
    name = NULL
  )
)

hadley3 <- Person$new("Hadley")
hadley3

hadley3$name

```
### 14.3.2 active

behave like fields but are defined as methods

### 14.3.3 Exercises

#### 1. Create a bank account class that prevents you from directly setting the account balance, but you can still withdraw from and deposit to. Throw an error if you attempt to go into overdraft.


```{r}
BankAccount <- R6Class("BankAcount", 
                       public = list(
                         initialize = function(amt = 0) {
                           private$balance = amt
                         },
                         deposit = function(amt) {
                           private$balance <- private$balance + amt
                           invisible(self)
                         },
                         withdraw = function(amt) {
                           private$balance <- private$balance - amt
                           invisible(self)
                         },
                         print = function() {
                           cat("The current balance is: ", private$balance, "\n")
                           invisible(self)
                         }
                       ),
                       private = list(balance = NULL)
)

BA <- BankAccount$new(125)

BA

BA$deposit(100)
BA$withdraw(33)
BA
```


```{r, error=TRUE}
BA$balance <- 200
BA

BA1$balance <- 200 # the one with public balance
BA1
```


#### 2. Create a class with a write-only $password field. It should have $check_password(password) method that returns TRUE or FALSE, but there should be no way to view the complete password.

```{r, error=TRUE}
authenticator <- R6Class("authenticator",
                         public = list(
                           initialize = function(passwd) {
                             private$.password <- passwd
                           },
                           print = function() {
                             cat("It is not possible to view the password.\n")
                             cat("Use `check_password()` method to check a password\n")
                           },
                           check_password = function(passwd) {
                             passwd == private$.password
                           }
                         ),
                         private = list(
                           .password = NULL),
                         active = list(
                           password = function(value) {
                             if(missing(value)) {
                               stop("password is write-only\n", call.=FALSE)
                            } else {
                              private$.password <- value
                             }
                           }
                         ))

au <- authenticator$new("ps1")
au

au$check_password("ps2")
au$check_password("ps1")
au$password
au$password <- "ps3"
au$check_password("ps3")
```


#### 3. Extend the Rando class with another active binding that allows you to access the previous random value. Ensure that active binding is the only way to access the value.

#### 4. Can subclasses access private fields/methods from their parent? Perform an experiment to find out.

### 14.4 Reference Semantics

R6 are reference semantics, and are not copied on modify!!

```{r}
y1 <- Accumulator$new() 
y2 <- y1

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)

```

Need to use clone

```{r}
y1 <- Accumulator$new() 
y2 <- y1$clone()

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)

```

