---
title: "Chapter 9"
author: "Julin Maloof"
date: "2023-01-16"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 9

```{r}
library(tidyverse)
```


## 9.2.6 Exercises

### 1. 
Use `as_mapper()` to explore how purrr generates anonymous functions for the integer, character, and list helpers. What helper allows you to extract attributes? Read the documentation to find out.

```{r}
f <- as_mapper(~ . + 1)
f
f(1:3)
```
```{r}
f <- as_mapper(2)
f
f(3:1)
```


```{r}
f <- as_mapper(c("a", "b", "c"))
f
f(1:3)
f(c(a=4,b=3, c=1))
f(list(a=4,b=3,c=1))
f(list(a=list(b=list(c=23))))
```
Ahhh, each argument takes you one level deeper into the list

Can use attr_getter

```{r}
f <- attr_getter("class")
f
f(mpg)
```


### 2.
map(1:3, ~ runif(2)) is a useful pattern for generating random numbers, but map(1:3, runif(2)) is not. Why not? Can you explain why it returns the result that it does?

```{r}
map(1:3, ~ runif(2))
```
```{r}
map(1:3, runif(2))
```
```{r}
as_mapper(runif(2))
```

_In the second form, the output from `runif(2)` is taken as input to `as_mapper` and interpreted as positions? names? to extract 


### 3.
Use the appropriate map() function to:

a. Compute the standard deviation of every column in a numeric data frame.

```{r}
df <- as_data_frame(matrix(rnorm(100),ncol=5))
df

map_dbl(df, sd)
```


b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to do it in two steps.)

```{r}
columns <- map_lgl(mpg, is.numeric)
map_dbl(mpg[columns], sd)
```
Or

```{r}
mpg %>%
  summarize(across(.cols = where(is.numeric), sd ))
```

c. Compute the number of levels for every factor in a data frame.


```{r}
columns <- map_lgl(iris, is.factor)
map_dbl(iris[columns], nlevels)
```

### 4.
The following code simulates the performance of a t-test for non-normal data. Extract the p-value from each test, then visualise.
```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))

map_dbl(trials, "p.value") %>% 
  hist()
```

### 5.
The following code uses a map nested inside another map to apply a function to every element of a nested list. Why does it fail, and what do you need to do to make it work?

```{r}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3

map(x, ~ map(.x, .f= triple))

#map(x, map, .f = triple)
#> Error in .f(.x[[i]], ...): unused argument (function (.x, .f, ...)
#> {
#> .f <- as_mapper(.f, ...)
#> .Call(map_impl, environment(), ".x", ".f", "list")
#> })
```
_The problem was that the ".f" argument was going to the outer map._

### 6.

Use map() to fit linear models to the mtcars dataset using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

map(formulas, lm, data=mtcars)
```

Better...

```{r}
tib <- tibble(formulas = list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
))

tib <- tib %>% 
  mutate(lm = map(formulas, lm, mtcars))

tib %>% mutate(glance = map(lm, broom::glance)) %>% unnest(glance)

tib %>% mutate(tidy = map(lm, broom::tidy)) %>% unnest(tidy)

```

### 7.

Fit the model mpg ~ disp to each of the bootstrap replicates of mtcars in the list below, then extract the $R^2$ of the model fit (Hint: you can compute the $R^2$ with summary().)

```{r}
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))

map(bootstraps, ~ lm(mpg ~ disp, .x)) %>%
  map(summary) %>%
  map_dbl("r.squared")
```

## 9.4.6 Exercises

### 1.
Explain the results of modify(mtcars, 1).

```{r}
mtcars
```

```{r}
modify(mtcars, 1)
```
_clearly it is taking the first row, but keeps the rownames.  I would have thought it would keep the first column._

_Oh, I see it is going to go through each item in the last (each column) and then take the first item there_


### 2.
Rewrite the following code to use iwalk() instead of walk2(). What are the advantages and disadvantages?

```{r}
temp <- "./temp"
cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(cyls, paths, write.csv)
```

```{r}
cyls <- split(mtcars, mtcars$cyl)
iwalk(cyls, ~ write.csv(.x, file.path(temp, paste0("cyl-", .y, ".csv"))))
```

_fewer lines and variable, but maybe harder to read_

### 3. Explain how the following code transforms a data frame using functions stored in a list.

```{r}
trans <- list(
  disp = function(x) x * 0.0163871,
  am = function(x) factor(x, labels = c("auto", "manual"))
)

nm <- names(trans)
mtcars[nm] <- map2(trans, mtcars[nm], function(f, var) f(var))
```

_trans is a list of 2 functions; mtcars[nm] is a list of 2 columns.  map applies the first function to the first column, etc_.

Compare and contrast the map2() approach to this map() approach:

```{r}
mtcars[nm] <- map(nm, ~ trans[[.x]](mtcars[[.x]]))
```

_ugly!_

### 4.
What does write.csv() return, i.e. what happens if you use it with map2() instead of walk2()?

```{r}
temp <- "./temp"
cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
map2(cyls, paths, write.csv)
```

_NULL_