---
title: "Chapter 14"
author: "Julin Maloof"
date: "2023-04-07"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R6

__"in most cases R6 will lead you to non-idiomatic R code."__

```{r}
library(R6)
```

Use the `R6Class()` function to define a class.  Assign it to an object that has the same name as the class.  First argument is class name, second argument is public fields and methods.

```{r}
Accumulator <- R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)
```

```{r}
Accumulator
```

Create a new object of this clas with $new function

```{r}
x <- Accumulator$new()
```

```{r}
x$sum
x$add(4)
x$sum
```

## 14.2.1 Method Chaining
 
side-effect R6 methods (those that update a field and are not called to create an output) should return self invisibly

then you can do things like this
```{r}
x$
  add(10)$
  add(10)$
  sum
```

## 14.2.2 Important methods

Generally should define $initialize and $print methods to override the defaults

```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) {
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    invisible(self)
  }
))

hadley2 <- Person$new("Hadley")
hadley2
```

## 14.2.3 Adding methods after creation

Use $set

```{r}
Accumulator <- R6Class("Accumulator")
Accumulator$set("public", "sum", 0)
Accumulator$set("public", "add", function(x = 1) {
  self$sum <- self$sum + x 
  invisible(self)
})
```

Will only impact newly created objects, will not be added to previosly created objects of the same class.

## 14.2.5 Inheritance

```{r}
AccumulatorChatty <- R6Class("AccumulatorChatty", 
  inherit = Accumulator,
  public = list(
    add = function(x = 1) {
      cat("Adding ", x, "\n", sep = "")
      super$add(x = x)
    }
  )
)

x2 <- AccumulatorChatty$new()
x2$add(10)$add(1)$sum
```

## 14.2.5 Introspections

Every R6 object has an S3 class hierarchy
```{r}
class(hadley2)
```

## 14.2.6. Exercises

### 1 Create a bank account R6 class that stores a balance and allows you to deposit and withdraw money. Create a subclass that throws an error if you attempt to go into overdraft. Create another subclass that allows you to go into overdraft, but charges you a fee.

```{r}
BankAccountBasic <- R6Class("BankAcountSimple", list(
  balance = 0,
  deposit = function(amt) {
    self$balance <- self$balance + amt
    invisible(self)
  },
  withdraw = function(amt) {
    self$balance <- self$balance - amt
    invisible(self)
  },
  print = function() {
    cat("The current balance is: ", self$balance, "\n")
    invisible(self)
  }
))

BA1 <- BankAccountBasic$new()

BA1$deposit(100)
BA1$withdraw(33)
BA1
```

```{r, error=TRUE}
BankAccount2 <- R6Class("BankAcount2",
                       inherit = BankAccountBasic,
                       public = list(
                         withdraw=function(amt) {
                           stopifnot(self$balance-amt>=0)
                           super$withdraw(amt)
                         }
                       ))

BA2 <- BankAccount2$new()

BA2$deposit(100)
BA2$withdraw(33)
BA2
BA2$withdraw(100)
BA2
```
```{r, error=TRUE}
BankAccount3 <- R6Class("BankAcount3",
                       inherit = BankAccountBasic,
                       public = list(
                         withdraw=function(amt) {
                           super$withdraw(amt)
                           if (self$balance < 0) {
                             warning("New balance is less than $0, a $5.00 fee is being assessed\n")
                             super$withdraw(5)
                           }
                         }
                       ))

BA3 <- BankAccount3$new()

BA3$deposit(100)
BA3$withdraw(150)
BA3
```

### 2. Create an R6 class that represents a shuffled deck of cards. You should be able to draw cards from the deck with $draw(n), and return all cards to the deck and reshuffle with $reshuffle(). Use the following code to make a vector of cards.

```{r}
deck <- R6Class("deck", public = list(
  stack=character(),
  drawn=character(),
  initialize = function() {
    suit <- c("♠", "♥", "♦", "♣")
    value <- c("A", 2:10, "J", "Q", "K")
    self$stack <- sample(paste0(rep(value, 4), suit))
    self$drawn <- character()
  },
  reshuffle = function() {
    self$stack <- c(self$stack, self$drawn)
    self$drawn <- character()
    self$stack <- sample(self$stack)
  },
  draw = function(n) {
    newdraw <- self$stack[1:n]
    self$drawn <- c(self$drawn, newdraw)
    self$stack <- self$stack[-1:-n]
    cat(newdraw)
  },
  print = function() {
    cat("Not Drawn: ",self$stack, "\n")
    cat("Drawn: ", self$drawn, "\n")
  }
))

mydeck <- deck$new()
mydeck

mydeck$draw(5)
mydeck

mydeck$draw(5)
mydeck

mydeck$reshuffle()
mydeck

```
### 3. Why can’t you model a bank account or a deck of cards with an S3 class?

Beacuse S3 is not modify in place

### 4. Create an R6 class that allows you to get and set the current time zone. You can access the current time zone with Sys.timezone() and set it with Sys.setenv(TZ = "newtimezone"). When setting the time zone, make sure the new time zone is in the list provided by OlsonNames().

```{r, error=TRUE}
TZ <- R6Class("TZ", public=list(
  tz=Sys.timezone(),
  get=function() self$tz,
  set=function(new_tz) {
    stopifnot(new_tz %in% OlsonNames())
    Sys.setenv(TZ = new_tz)
    self$tz <- Sys.timezone()
  }
))

tz <- TZ$new()
tz$get()
tz$set("US/Hawaii") # I wish
tz$get()
tz$set("Mars")
tz$get()
```

### 5. Create an R6 class that manages the current working directory. It should have $get() and $set() methods.

```{r, error=TRUE}
WD <- R6Class("WD", public=list(
  wd=getwd(),
  get=function() self$wd,
  set=function(new_wd) {
    stopifnot(dir.exists(new_wd))
    setwd(new_wd)
    self$wd <- new_wd
  }
))

wd <- WD$new()
wd$get()
wd$set("../") 
wd$get()
wd$set("Mars")
wd$get()
wd$set("/Users/jmaloof/git/AdvancedR")
wd$get()
```

### 5. Why can’t you model the time zone or current working directory with an S3 class?

no modify in place?

### 6. What base type are R6 objects built on top of? What attributes do they have?