---
title: "Chapter 19"
author: "Julin Maloof"
date: "2023-07-20"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rlang)
library(purrr)
```


# Quasiquotation

## 19.1

quotation is the act of capturing an unevaluated expression

unquotation is the ability to selectively unquote parts of a quoted expression.

Together these are _quasiquoation_.  Quasiquotation enables combining code inherent to the function with that provided by the user.

## 19.2 Motivation

This is a pain:

```{r}
paste("Good", "morning", "Hadley")
paste("Good", "afternoon", "Alice")
```

Nicer to do it without typing all of the quotes

```{r}
cement <- function(...) {
  args <- ensyms(...)
  paste(purrr::map(args, as_string), collapse = " ")
}

cement(Good, morning, Hadley)
#> [1] "Good morning Hadley"
cement(Good, afternoon, Alice)
#> [1] "Good afternoon Alice"
```

But what if we want to do it with objects? 

```{r}
#works with paste:
name <- "Hadley"
time <- "morning"

paste("Good", time, name)

#but not with cement:
cement(Good, time, name)

```

We can solve this by using !! to drop the implicit quotes:

```{r}
cement(Good, !!time, !!name)
```

Comparing cement and paste.  paste we have to quote literals.  Cement we have to unquote variable.

```{r}
paste("Good", time, name)
cement(Good, !!time, !!name)
```

### 19.2.1 Vocabulary

Distinction between evaluated and quoted arguements.

_evaluated_ follows R's normal evaluation rules
_quoted_ are captured expressions that are treated in some custom way.

paste evaluates all of its arguements, whereas cement quotes all of its arguments.

### 19.2.2 Exercises

#### 1. For each function, identify which arguments are quoted and which are evaluated

```{r}
library(MASS) # quoted

mtcars2 <- subset(mtcars, cyl == 4) # first evaluated, second quoted

with(mtcars2, sum(vs)) #first is evaluated, second is quoted
sum(mtcars2$am) #evaluated

rm(mtcars2) #evaluated(?)
```

#### 2. For each function in the following tidyverse code, identify which arguments are quoted and which are evaluated.

```{r}
library(dplyr) # quoted
library(ggplot2) # quoted

by_cyl <- mtcars %>%
  group_by(cyl) %>% #quoted
  summarise(mean = mean(mpg)) #quoted

ggplot(by_cyl, aes(cyl, mean)) + geom_point() # by_cyl evaluated, cyl, mean quoted
```


## 19.3 Quoting

Capturing an expression without evaluating it

Need a pair of functions because the expression can be supplied directly or indirectly via lazily-evaluated function argument.  (Hopefully there is an example...)

### 19.3.1 Capturing expressions

There are four important quoting functions. For interactive use, `expr()` is the most important. It captures its argument exactly as provided.

```{r}
expr(x + y)

expr(1 / 2 / 3)
```

But not great inside a function because it captures what the function developer typed.

```{r}
f1 <- function(x) expr(x)
f1(a + b + c)
#> x
```

To solve this problem we use another function, `enexpr()`.  This captures what the caller supplied to the. function.

```{r}
f2 <- function(x) enexpr(x)
f2(a + b + c)
```

expr and exprs for capturing expressions we type

enexpr and enexprs for capturing expressions that the function caller types

### 19.3.2 capturing symbols

`ensym` and `ensyms`: similar to `enexpr`, but through an error if given anything byt a symbol (or string)

```{r, error=TRUE}
f <- function(...) ensyms(...)
f(x)

f("x")


f(2+3)
```

### 19.3.3 with base R

The base R functions do not support unquoting

base equivalent of `expr` is `quote`

base equivalent of `enexpr` is `substitute`

base equivalent of `exprs` is `alist`

also note that `~` is a quoting function that captures the environment.

### 19.3.4 substitution

in addition to quoting, `substitute` will substitute in the values of symbols in the current environment

### 19.3.6 Exercises:

#### 1. how is expr() implemented?

```{r}
expr
```

#### 2 Compare and contrast the following two functions. Can you predict the output before running them?

```{r}
f1 <- function(x, y) {
  exprs(x = x, y = y)
}
f2 <- function(x, y) {
  enexprs(x = x, y = y)
}
f1(a + b, c + d) # --> "x=x, y=y"
f2(a + b, c + d) # -->
#[x]
#a + b
#[y]
#c+d

```

#### 3 What happens if you try to use enexpr() with an expression (i.e.  enexpr(x + y) ? What happens if enexpr() is passed a missing argument?

```{r, error=TRUE}
enexpr(x + y)
enexpr(NA)
```

Throws errors

#### 4. How are exprs(a) and exprs(a = ) different? Think about both the input and the output.

First one should capture "a".  Second one is setting "a" as the name for the first item in the list.

```{r}
exprs(a)
exprs(a=c+d)
```

#### 5. What are other differences between exprs() and alist()? Read the documentation for the named arguments of exprs() to find out.

```{r, eval=FALSE}
?exprs
```

options for whether or not names are assigned to the unamed arguments, what to do with empty arguments, and something about unquoting names that I don't understand.

#### 6. The documentation for substitute() says:

Substitution takes place by examining each component of the parse tree as follows:

* If it is not a bound symbol in env, it is unchanged.
* If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol.
* If it is an ordinary variable, its value is substituted, unless env is .GlobalEnv in which case the symbol is left unchanged.

Create examples that illustrate each of the above cases.

```{r}
# 1
x <- 5
substitute(x+y*2, list(y=4))
#x not substituted, y substituted

substitute(x+y*2, .GlobalEnv)

substitute(x+y*2, list(y=4, x=x))

substitute(x+y*2, as.list(.GlobalEnv))

```


```{r}
#2
f6_2 <- function(x)
  substitute(x+y*2)
f6_2(x=a)
```


```{r}
#3
f6_3 <- function(y) {
  x <- 5 
  substitute(x+y*2)
}
f6_3(y=4)
#compare with # 1

f6_3 <- function(y) {
  x <- 5 
  substitute(x+y*2, .GlobalEnv)
}
f6_3(y=4)
```

## 19.4

### 19.4.1 unquote one argument

Use `!!`

### 19.4.2 unquoting a function

You can also use `!!` but pay attention to operatior precedence.  You need an extra set of parentheses

```{r, error=TRUE}
# correct way to only unquote f
x <- 1
y <- 2
f <- expr(foo)
expr((!!f)(x, y))
#> foo(x, y)
#> 
#> #compare to
 expr(!!f(x,y)) # error!
```

Works when f is a call but perhaps cleaner to use rlang::call2

```{r}
f <- expr(pkg::foo)
expr((!!f)(x, y))
#> pkg::foo(x, y)


call2(f, expr(x), expr(y))

```

### 19.4.3 unquoting a missing argument

```{r, error=TRUE}
arg <- missing_arg()
expr(foo(!!arg, !!arg))
#> Error in enexpr(expr): argument "arg" is missing, with no default
```

Instead
```{r}
expr(foo(!!maybe_missing(arg), !!maybe_missing(arg)))
```

### 19.4.4 Unquoting in special forms

infix function `$` causes problems.  Workaround is to use the pre-fix version:

```{r}
x <- expr(x)
expr(`$`(df, !!x))
#> df$x
```

### 19.4.5 Unquoting many arguments

Use `!!!`, takes a list of expressions, pastes them together and inserts them at the !!!

```{r}
xs <- exprs(1, a, -b)
expr(f(!!!xs, y))
#> f(1, a, -b, y)

# Or with names
ys <- set_names(xs, c("a", "b", "c"))
expr(f(!!!ys, d = 4))
#> f(a = 1, b = a, c = -b, d = 4)
```

`!!!` can be used anywhere that `...` is accepted.  Helpful in `call2`

```{r}
call2("f", !!!xs, expr(y))
#> f(1, a, -b, y)
```

### 19.4.6 fiction of !!

These arent' real operators.  Can only use in Rlang aware situations

### 19.4.7 misleading ASTs

beware inline functions.  expr_print or lobstr::ast can help resolve confusion.

### 19.4.8 Exercises

Given the following components:

```{r}
xy <- expr(x + y)
xz <- expr(x + z)
yz <- expr(y + z)
abc <- exprs(a, b, c)
```

Use quasiquotation to construct the following calls:


```{r}
(x + y) / (y + z)
-(x + z) ^ (y + z)
(x + y) + (y + z) - (x + y)
atan2(x + y, y + z)
sum(x + y, x + y, y + z)
sum(a, b, c)
mean(c(a, b, c), na.rm = TRUE)
foo(a = x + y, b = y + z)
```

```{r}
expr(!!xy /!!xz)

expr(-(!!xy)  ^ (!!xz))

expr((!!(xy)) + !!yz - !!xy) # how to get parentheses?

expr(atan2(!!xy, !!yz))

expr(sum(!!xy, !!xy, !!yz))

expr(sum(!!!abc))

expr(mean(c(!!!abc), na.rm = TRUE)) #was I supposed to do this some more complicated way?

expr(foo(a=!!xy, b=!!yz))
```


#### 2 The following two calls print the same, but are actually different:

```{r}
(a <- expr(mean(1:10)))
#> mean(1:10)
(b <- expr(mean(!!(1:10))))
#> mean(1:10)
identical(a, b)
#> [1] FALSE
```

What is the difference and which is more natural?

b is going to go ahead and evaluate 1:10, while a won't.  Why not use a?

```{r}
expr_print(a)

expr_print(b)
```

## 19.5 Non-quoting

Base R mostly selectively turns off quoting as needed, rather than unquoting.  This is called non-quoting.

## 19.6 ...

### 19.6 Exercises

#### 1. One way to implement exec() is shown below. Describe how it works. What are the key ideas?

```{r}
exec <- function(f, ..., .env = caller_env()) {
  args <- list2(...)
  do.call(f, args, envir = .env)
}
```

Captures the calling environment, but allows user to set something else

uses `list2` to deal with the possibility that `...` is a list or not.

#### 2. Carefully read the source code for interaction(), expand.grid(), and par(). Compare and contrast the techniques they use for switching between dots and list behaviour.

```{r}
interaction
```

```{r}
expand.grid
```
```{r}
par
```

interaction checks to see if the first item in ... is a list, and if so it uses that.  expand.grid seems functionally equivalent

#### 3 Explain the problem with this definition of set_attr()

```{r, error=TRUE}
set_attr <- function(x, ...) {
  attr <- rlang::list2(...)
  #print(attr)
  attributes(x) <- attr
  x
}
set_attr(1:10, x = 10)
#> Error in attributes(x) <- attr: attributes must be named
```
Problem is that we have a named argument `x`.  When we call the function, x ends up being 10 and ... is the (unamed) vector `1:10`
```{r, error=TRUE}
set_attr <- function(x, ...) {
  attr <- rlang::list2(..., .named=TRUE)
  attributes(y) <- attr
  x
}
set_attr(1:10, y = 10)
```

### 14.7 exercises

#### 1. In the linear-model example, we could replace the expr() in reduce(summands, ~ expr(!!.x + !!.y)) with call2(): reduce(summands, call2, "+"). Compare and contrast the two approaches. Which do you think is easier to read?

```{r}
linear <- function(var, val) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)

  reduce(summands, ~ expr(!!.x + !!.y))
}

linear(x, c(10, 5, -4))
```

```{r, error=TRUE}
linear <- function(var, val) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)

  reduce(summands, call2, "+")
}

linear(x, c(10, 5, -4))
```

#### 2. Re-implement the Box-Cox transform defined below using unquoting and new_function():

```{r}
bc <- function(lambda) {
  if (lambda == 0) {
    function(x) log(x)
  } else {
    function(x) (x ^ lambda - 1) / lambda
  }
}

bc(0)

bc(0.5)
```

```{r}
bc2 <- function(lambda) {
  if (lambda == 0) {
    new_function(exprs(x=), expr(log(x) ))
  } else {
    new_function(exprs(x=), expr((x ^ !!lambda - 1) / !!lambda))
  }
}

bc2(0)

bc2(0.5)
```
#### 3 Re-implement the simple compose() defined below using quasiquotation and new_function():

```{r}
compose <- function(f, g) {
  function(...) f(g(...))
}

compose(mean, median)
```


```{r}
compose2 <- function(f, g) {
  f <- enexpr(f)
  g <- enexpr(g)
  new_function(exprs(...=), expr((!!f)((!!g)(...))))
}

compose2(mean, median)
```

