---
title: "Chapter 17"
author: "Julin Maloof"
date: "2023-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 17 Metaprogramming overview

```{r}
library(rlang)
library(lobstr)
```

## 17.2 Code is data

can use `rlang::expr()` to capture code:

```{r}
expr(mean(x, na.rm = TRUE))
#> mean(x, na.rm = TRUE)
expr(10 + 100 + 1000)
#> 10 + 100 + 1000
```

Captures code is called an `expression`

`expr()` Doesn't work in arguments passed to functions.  but `enexpr()` does.  This quotes the argument

```{r}
capture_it <- function(x) {
  enexpr(x)
}
capture_it(a + b + c)
#> a + b + c
```


Captured code behaves like a list and can be modified:

```{r}
f <- expr(f(x = 1, y = 2))

# Add a new argument
f$z <- 3
f

# Or remove an argument:
f[[2]] <- NULL
f
```

Note that the first element is the function to be called

```{r}
f[[1]]
```

## 17.3 Code as a tree

Code can be displayed as an abstract syntax tree (AST)

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
```

## 17.4 Code cqn generate code

`call2()` can generate a function call with arguments:

```{r}
call2("f", 1, 2, 3)
call2("+", 1, call2("*", 2, 3))
```


You can also use `expr` and `enexpr` and the unquote `!!` operator

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

xx
yy
```


```{r}
expr(xx / yy)

expr(!!xx / !!yy)
```

Can build up functions from user input (although I don't really understand this example)

```{r}
cv <- function(var) {
  var <- enexpr(var)
  expr(sd(!!var) / mean(!!var))
}

cv(x)
#> sd(x)/mean(x)
cv(x + y)
#> sd(x + y)/mean(x + y)
```

## 17.5 evaluation runs code

`eval` runs an expression.  You can give it an environment or it will use the current environment.

```{r}
eval(expr(x + y), env(x = 1, y = 10))
#> [1] 11
eval(expr(x + y), env(x = 2, y = 100))
#> [1] 102
```


## 17.6 Customizing evaluation of functions

We can overide the meaning of functions in a given environment

```{r}
string_math <- function(x) {
  e <- env(
    caller_env(),
    `+` = function(x, y) paste0(x, y),
    `*` = function(x, y) strrep(x, y)
  )

  eval(enexpr(x), e)
}

name <- "Hadley"
string_math("Hello " + name)
#> [1] "Hello Hadley"
string_math(("x" * 2 + "-y") * 3)
#> [1] "xx-yxx-yxx-y"
```

## 17.7 Customizing evaluation with data

provide a data mask using `eval_tidy`

```{r}
df <- data.frame(x = 1:5, y = sample(5))
eval_tidy(expr(x + y), df)
#> [1] 6 6 4 6 8
```

Can create a second `with`

```{r}
with2 <- function(df, expr) {
  eval_tidy(enexpr(expr), df)
}

with2(df, x + y)
#> [1] 6 6 4 6 8
```

but there is a bug...

## 17.8 quosures

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}
```

```{r}
df <- data.frame(x = 1:3)
a <- 10
with2(df, x + a)
#> [1] 1001 1002 1003
```

using the wrong `a`!

Solve by using an quosure.  This bundles the expresion with its environment

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}

with2(df, x + a)
#> [1] 11 12 13
```


