---
title: "Chapter6"
author: "Julin Maloof"
date: "2022-10-08"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Quiz

1. What are the three components of a function?

_name, arguments, code?_

2. What does the following code return?

```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```
_11.  Note that the initial assignment of x <- 10 is just a distractor.  _


3. How would you usually write this code?

```{r}
`+`(1, `*`(2, 3))
```

```{r}
1 + 2*3
```


4. How could you make this call easier to read?

mean(, TRUE, x = c(1:10, NA))

```{r}
mean(, TRUE, x = c(1:10, NA))
```

```{r}
mean(c(1:10, NA), na.rm=TRUE)
```


5. Does the following code throw an error when executed? Why or why not?

```{r}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

I guessed "no" and I was correct.  Because "b" is not evaluated

6. What is an infix function? How do you write it? What’s a replacement function? How do you write it?

??

7. How do you ensure that cleanup action occurs regardless of how a function exits?

??

## Exercises 6.2.5

1. Given a name, like "mean", match.fun() lets you find a function. Given a function, can you find its name? Why doesn’t that make sense in R?

_I don't even know what this is asking._

2. It’s possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?

```{r}
function(x) 3()
#> function(x) 3()
(function(x) 3)()
#> [1] 3
```

_Took me a minute to understand this at all.  But the function is a function that will return 3, no matter what is input.  The second veresion is what works but I do not understand the syntax_


3. A good rule of thumb is that an anonymous function should fit on one line and shouldn’t need to use {}. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?

4. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
f <- function(x) x^2

typeof(f)

is.function(f)

is.primitive(f)

is.primitive(sum)
```


5. This code makes a list of all functions in the base package.

```{r}
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Use it to answer the following questions:

6. Which base function has the most arguments?

```{r}
fun.args <- sapply(funs, function(x) length(formals(x))) 

names(funs)[which.max(fun.args)] 

formals(scan)
```


7. How many base functions have no arguments? What’s special about those functions?

```{r}
names(funs)[fun.args==0]

sapply(funs[fun.args==0], is.primitive)
```

_many are primitives, the rest return something without input_


8. How could you adapt the code to find all primitive functions?

```{r}
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.function, objs)
primfuns <- Filter(is.primitive, objs)
```

9. What are the three important components of a function?

_Arguments, body, environment_

10. When does printing a function not show the environment it was created in?

_when the function exists in the global environment_

```{r}
sum
f
stats::aggregate
```

## 6.4.5 Exercises

1. What does the following code return? Why? Describe how each of the three c’s is interpreted.

```{r}
c <- 10
c(c = c)
```

_I predict 10.  Answer is c=10.  The first "c" is an object name, a reference to the value 10.  The second c calls the `c()` function.  The third c is used as a name for the first element in the vector that c is creating, and the final c refers to the "c" object._

2. What are the four principles that govern how R looks for values?

_masking, fresh start, functions separate from variables, dynamic lookup_

3. What does the following function return? Make a prediction before running the code yourself.

```{r}
f <- function(x) { # this is the f that gets called
  # x = 10
  f <- function(x) {
    f <- function() {
      x ^ 2
    }
    f() + 1 
  }
  f(x) * 2 # calls the first nested function, which calls the innermost function.  So 202
}
f(10)
```

## 6.5.4 Exercises

1. What important property of && makes x_ok() work?

```{r}
x_ok <- function(x) {
  !is.null(x) && length(x) == 1 && x > 0
}

x_ok(NULL)
#> [1] FALSE
x_ok(1)
#> [1] TRUE
x_ok(1:3)
#> [1] FALSE
```

with `&&` it doesn't keep evaluating once a FALSE is reached

What is different with this code? Why is this behaviour undesirable here?

```{r}

x_ok <- function(x) {
  !is.null(x) & length(x) == 1 & x > 0
}

x_ok(NULL)
#> logical(0)
x_ok(1)
#> [1] TRUE
x_ok(1:3)
#> [1] FALSE FALSE FALSE
```

`&` is going to evaluate all code.  We want a single T or F and that is not what is being returned here.

2. What does this function return? Why? Which principle does it illustrate?

```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

_should return 100 and illustrates that arguments can be assigned internal in the function.  I hate this_


3. What does this function return? Why? Which principle does it illustrate?

```{r}
y <- 10
f1 <- function(x = {y <- 1; 2}, y = 0) {
  c(x, y)
}
f1()
y
```
_I don't get it_

4. In hist(), the default value of xlim is range(breaks), the default value for breaks is "Sturges", and

```{r}
range("Sturges")
#> [1] "Sturges" "Sturges"
```


Explain how hist() works to get a correct xlim value.

_it is using the Sturges function_

5. Explain why this function works. Why is it confusing?

```{r}
show_time <- function(x = stop("Error!")) {
  stop <- function(...) Sys.time()
  print(x)
}
show_time()
#> [1] "2021-02-21 19:22:36 UTC"
```

_because x is evaluated inside the function and a new stop function has been defined inside of the show_time function_

6. How many arguments are required when calling library()?

```{r}
args(library)
#library()
```
Zero arguments, but not sure how to figure this out.

## 6.6.1 Exercises
### 1. Explain the following results:

```{r}
sum(1, 2, 3)
#> [1] 6
mean(1, 2, 3)
#> [1] 1

sum(1, 2, 3, na.omit = TRUE)
#> [1] 7
mean(1, 2, 3, na.omit = TRUE)
#> [1] 1
```


_Part 1: `sum()` has a `...` argument first and it will sum up all argumetns given, aside from its final argument `na.rm = FALSE`.  In `mean()`, the `...` argument comes last; `mean()` only takes the mean of its first argument; the `...` argument is for passing arguments to other functions._

_Part 2: Somehow `na.omit = TRUE` is getting evaluated to `1` in `sum` and included in the total.  This doesn't happen for mean for the reasons described above_


### 2. Explain how to find the documentation for the named arguments in the following function call:

```{r}
plot(1:10, col = "red", pch = 20, xlab = "x", col.lab = "blue")
```

_`?plot` tells us that `...` is passed to other methods, such as `par()` so look at `?par`_

### 3. Why does plot(1:10, col = "red") only colour the points, not the axes or labels? Read the source code of plot.default() to find out.

_Because `plot.default` uses `Axis` but when it calls `Axis` is excludes `col` from the `...` that gets passed to `Axis`.  And labels color has to be specified with col.lab_

Test:

```{r}
plot.default(1:10, col="red")
```

```{r}
plot.julin <- function(x, y = NULL, type = "p", xlim = NULL, ylim = NULL, 
                       log = "", main = NULL, sub = NULL, xlab = NULL, ylab = NULL, 
                       ann = par("ann"), axes = TRUE, frame.plot = axes, panel.first = NULL, 
                       panel.last = NULL, asp = NA, xgap.axis = NA, ygap.axis = NA, 
                       ...) 
{
  # localAxis <- function(..., col, bg, pch, cex, lty, lwd) Axis(...)
  localAxis <- function(..., bg, pch, cex, lty, lwd) Axis(...)
  localBox <- function(..., col, bg, pch, cex, lty, lwd) box(...)
  localWindow <- function(..., col, bg, pch, cex, lty, lwd) plot.window(...)
  #  localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(...)
  localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(col.lab=col, ...)
  xlabel <- if (!missing(x)) 
    deparse1(substitute(x))
  ylabel <- if (!missing(y)) 
    deparse1(substitute(y))
  xy <- xy.coords(x, y, xlabel, ylabel, log)
  xlab <- if (is.null(xlab)) 
    xy$xlab
  else xlab
  ylab <- if (is.null(ylab)) 
    xy$ylab
  else ylab
  xlim <- if (is.null(xlim)) 
    range(xy$x[is.finite(xy$x)])
  else xlim
  ylim <- if (is.null(ylim)) 
    range(xy$y[is.finite(xy$y)])
  else ylim
  dev.hold()
  on.exit(dev.flush())
  plot.new()
  localWindow(xlim, ylim, log, asp, ...)
  panel.first
  plot.xy(xy, type, ...)
  panel.last
  if (axes) {
    localAxis(if (is.null(y)) 
      xy$x
      else x, side = 1, gap.axis = xgap.axis, ...)
    localAxis(if (is.null(y)) 
      x
      else y, side = 2, gap.axis = ygap.axis, ...)
  }
  if (frame.plot) 
    localBox(...)
  if (ann) 
    localTitle(main = main, sub = sub, xlab = xlab, ylab = ylab, 
               ...)
  invisible()
}
```

```{r}
plot.julin(1:10, col="red")

```

## 6.7.5 Exercises

### 1. What does load() return? Why don’t you normally see these values?

```{r}
save(list=ls(), file="test.Rdata")
withVisible(load("test.Rdata"))
file.remove("test.Rdata")
```
_`load()` returns (invisibly) the names of the items it has loaded_

### 2. What does write.table() return? What would be more useful?

```{r}
withVisible(
  write.table(matrix(1:25,ncol=5), file = "test.table")
)
file.remove("test.table")
```
_NULL. it might be more useful to return success or failure?_

### 3. How does the chdir parameter of source() compare to with_dir()? Why might you prefer one to the other?

```{r}
?source

?withr::with_dir
```

_looking at the code, `source` does more error checking of the path.  But I think I am missing something_

### 4. Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code works).

```{r}
gr.test <- function(code) {
  jpeg() # open graphics device
  on.exit(dev.off(), add=TRUE)
  
  force(code) # execute the code
}
```

```{r}
gr.test(plot(1:10, 1:10))

#gr.test(plat(1:10,1:10))

file.remove("Rplot001.jpeg")
```


### 5. We can use on.exit() to implement a simple version of capture.output().

```{r}
capture.output2 <- function(code) {
  temp <- tempfile()
  on.exit(file.remove(temp), add = TRUE, after = TRUE)

  sink(temp)
  on.exit(sink(), add = TRUE, after = TRUE)

  force(code)
  readLines(temp)
}
capture.output2(cat("a", "b", "c", sep = "\n"))
#> [1] "a" "b" "c"
```


Compare capture.output() to capture.output2(). How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas so they’re easier to understand?

_features removed: ability to name capture file, ability to append to file, ability to (not) split, ability to specify type_

_clarity: new function doesn't use a for loop, doesn't use ...elt()_

## 6.8.6 Exercises

### 1. Rewrite the following code snippets into prefix form:

```{r}
1 + 2 + 3

`+`(`+`(1,2), 3)

1 + (2 + 3)

`+`(1, `+`(2,3) )

x <- 1:10
n <- 7

if (length(x) <= 5) x[[5]] else x[[n]]

`if`( 
  `<=`(length(x), 5), #cond
  `[[`(x,5), #true
  `[[`(x,n) #false
)
  

```


### 2. Clarify the following list of odd function calls:

```{r}
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
x <- sample(c(1:10, NA), size = 20, replace = TRUE)

y <- runif(min = 0, max = 1, 20)
y <- runif(20, min = 0, max=1)

cor(m = "k", y = y, u = "p", x = x)
cor(x=x, y=y, use = "pairwise.complete.obs", method="kendall")

```


### 3. Explain why the following code fails:

```{r}
#modify(get("x"), 1) <- 10
#> Error: target of assignment expands to non-language object

```
_not sure what he is geting at here. Am I supposed to be using his modify function first? If so, it fails because `get("x")` gets the contents of x, not the object itself_

```{r}
`modify<-` <- function(x, position, value) {
  x[position] <- value
  x
}

x
get("x")

modify(x, 1) <- 10

```


### 4. Create a replacement function that modifies a random location in a vector.

```{r}
`modify.random<-` <- function(x, value) {
  position <- sample(length(x), size = 1)
  x[position] <- value
  x
}

z <- 1:10

modify.random(z) <- 200
z
```


### 5. Write your own version of + that pastes its inputs together if they are character vectors but behaves as usual otherwise. In other words, make this code work:

```{r}

`+` <- function(x,y) {
  if(any(is.character(c(x,y)))) 
    paste0(x,y)
  else
    sum(x,y)
}

1 + 2
#> [1] 3

"a" + "b"
#> [1] "ab"
```


### 6. Create a list of all the replacement functions found in the base package. Which ones are primitive functions? (Hint: use apropos().)

```{r}
replacement_functions <- apropos("<-$", mode = "function")

is_replacement_function <- sapply(replacement_functions, function(x) is.primitive(get(x)))

replacement_functions[is_replacement_function]
```


### 7. What are valid names for user-created infix functions?

_anything but "%"_

### 8. Create an infix xor() operator.

XOR is A or B but not both
```{r}
`%xor%` <- function(a,b) {
  (a|b) &! (a&b)
}

TRUE %xor% FALSE
FALSE %xor% TRUE
FALSE %xor% FALSE
TRUE %xor% TRUE

```


### 9. Create infix versions of the set functions intersect(), union(), and setdiff(). You might call them %n%, %u%, and %/% to match conventions from mathematics.

```{r}
`%n%` <- function(x,y) intersect(x,y)

`%u%` <- function(x,y) union(x,y)

`%/%` <- function(x,y) setdiff(x,y)
```

```{r}
(x <- c(sort(sample(1:20, 9)), NA))
(y <- c(sort(sample(3:23, 7)), NA))

union(x, y)
x %u% y

intersect(x, y)
x %n% y


setdiff(x, y)
x %/% y

setdiff(y, x)
y %/% x

```

