---
title: "Chapter 7"
author: "Julin Maloof"
date: "2022-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rlang)
```

What is the difference between `rlang::env()` and `new.env()`?

```{r}
e1 <- env(a = "John",
          b = 23,
          c = FALSE)
env_print(e1)
```

```{r}
e2 <- new.env()
e2$a <- "Paulo"
e2$b <- 42
e2$c <- TRUE

env_print(e2)
```

```{r}
env_parent() #why rlang?
```

```{r}
env_parents()
```

## 7.2.7 Exercises

### 1. List three ways in which an environment differs from a list.

_1. not ordered_
_2. names must be unique_
_3. has a parent_

### 2. Create an environment as illustrated by this picture.

```{r}
e2 <- env()
e2$loop <- e2
env_print(e2)
```


### 3. Create a pair of environments as illustrated by this picture.

```{r}
e3a <- env()
e3b <- env()
e3a$loop <- e3b
e3b$dedoop <- e3a

env_print(e3a)
env_print(e3b)
```


### 4. Explain why e[[1]] and e[c("a", "b")] don’t make sense when e is an environment.

_`e[[1]]` does not make sense because objects in an environment are not ordered so there is no "first" element_

_I am not as sure about `e[c("a", "b")]` but I guess because you can't subset an environment (but why not?)_

### 5. Create a version of env_poke() that will only bind new names, never re-bind old names. Some programming languages only do this, and are known as single assignment languages.



### 6. What does this function do? How does it differ from <<- and why might you prefer it?

```{r}
rebind <- function(name, value, env = caller_env()) {
  if (identical(env, empty_env())) {
    stop("Can't find `", name, "`", call. = FALSE)
  } else if (env_has(env, name)) {
    env_poke(env, name, value)
  } else {
    rebind(name, value, env_parent(env))
  }
}
rebind("a", 10)
#> Error: Can't find `a`
a <- 5
rebind("a", 10)
a
#> [1] 10
```

