---
title: "Chapter 19"
author: "Julin Maloof"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rlang)
library(purrr)
```


# Quasiquotation

## 19.1

quotation is the act of capturing an unevaluated expression

unquotation is the ability to selectively unquote parts of a quoted expression.

Together these are _quasiquoation_.  Quasiquotation enables combining code inherent to the function with that provided by the user.

## 19.2 Motivation

This is a pain:

```{r}
paste("Good", "morning", "Hadley")
paste("Good", "afternoon", "Alice")
```

Nicer to do it without typing all of the quotes

```{r}
cement <- function(...) {
  args <- ensyms(...)
  paste(purrr::map(args, as_string), collapse = " ")
}

cement(Good, morning, Hadley)
#> [1] "Good morning Hadley"
cement(Good, afternoon, Alice)
#> [1] "Good afternoon Alice"
```

But what if we want to do it with objects? 

```{r}
#works with paste:
name <- "Hadley"
time <- "morning"

paste("Good", time, name)

#but not with cement:
cement(Good, time, name)

```

We can solve this by using !! to drop the implicit quotes:

```{r}
cement(Good, !!time, !!name)
```

Comparing cement and paste.  paste we have to quote literals.  Cement we have to unquote variable.

```{r}
paste("Good", time, name)
cement(Good, !!time, !!name)
```

### 19.2.1 Vocabulary

Distinction between evaluated and quoted arguements.

_evaluated_ follows R's normal evaluation rules
_quoted_ are captured expressions that are treated in some custom way.

paste evaluates all of its arguements, whereas cement quotes all of its arguments.

### 19.2.2 Exercises

#### 1. For each function, identify which arguments are quoted and which are evaluated

```{r}
library(MASS) # quoted

mtcars2 <- subset(mtcars, cyl == 4) # first evaluated, second quoted

with(mtcars2, sum(vs)) #first is evaluated, second is quoted
sum(mtcars2$am) #evaluated

rm(mtcars2) #evaluated(?)
```

#### 2. For each function in the following tidyverse code, identify which arguments are quoted and which are evaluated.

```{r}
library(dplyr) # quoted
library(ggplot2) # quoted

by_cyl <- mtcars %>%
  group_by(cyl) %>% #quoted
  summarise(mean = mean(mpg)) #quoted

ggplot(by_cyl, aes(cyl, mean)) + geom_point() # by_cyl evaluated, cyl, mean quoted
```


## 19.3 Quoting

Capturing an expression without evaluating it

Need a pair of functions because the expression can be supplied directly or indirectly via lazily-evaluated function argument.  (Hopefully there is an example...)

### 19.3.1 Capturing expressions

There are four important quoting functions. For interactive use, `expr()` is the most important. It captures its argument exactly as provided.

```{r}
expr(x + y)

expr(1 / 2 / 3)
```

But not great inside a function because it captures what the function developer typed.

```{r}
f1 <- function(x) expr(x)
f1(a + b + c)
#> x
```

To solve this problem we use another function, `enexpr()`.  This captures what the caller supplied to the. function.

```{r}
f2 <- function(x) enexpr(x)
f2(a + b + c)
```

expr and exprs for capturing expressions we type

enexpr and enexprs for capturing expressions that the function caller types

### 19.3.2 capturing symbols

`ensym` and `ensyms`: similar to `enexpr`, but through an error if given anything byt a symbol (or string)

```{r, error=TRUE}
f <- function(...) ensyms(...)
f(x)

f("x")


f(2+3)
```

### 19.3.3 with base R

The base R functions do not support unquoting

base equivalent of `expr` is `quote`

base equivalent of `enexpr` is `substitute`

base equivalent of `exprs` is `alist`

also note that `~` is a quoting function that captures the environment.

### 19.3.4 substitution

in addition to quoting, `substitute` will substitute in the values of symbols in the current environment

### 18.3.6 Exercises:

#### 1. how is expr() implemented?

```{r}
expr
```

#### 2 Compare and contrast the following two functions. Can you predict the output before running them?

```{r}
f1 <- function(x, y) {
  exprs(x = x, y = y)
}
f2 <- function(x, y) {
  enexprs(x = x, y = y)
}
f1(a + b, c + d) # --> "x=x, y=y"
f2(a + b, c + d) # -->
#[x]
#a + b
#[y]
#c+d

```

#### 3 What happens if you try to use enexpr() with an expression (i.e.  enexpr(x + y) ? What happens if enexpr() is passed a missing argument?

```{r, error=TRUE}
enexpr(x + y)
enexpr(NA)
```

Throws errors

#### 4. How are exprs(a) and exprs(a = ) different? Think about both the input and the output.

First one should capture "a".  Second one is setting "a" as the name for the first item in the list.

```{r}
exprs(a)
exprs(a=c+d)
```

#### 5. What are other differences between exprs() and alist()? Read the documentation for the named arguments of exprs() to find out.

```{r, eval=FALSE}
?exprs
```

options for whether or not names are assigned to the unamed arguments, what to do with empty arguments, and something about unquoting names that I don't understand.

#### 6. The documentation for substitute() says:

Substitution takes place by examining each component of the parse tree as follows:

* If it is not a bound symbol in env, it is unchanged.
* If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol.
* If it is an ordinary variable, its value is substituted, unless env is .GlobalEnv in which case the symbol is left unchanged.

Create examples that illustrate each of the above cases.

```{r}
# 1
x <- 5
substitute(x+y*2, list(y=4))
#x not substituted, y substituted
```


```{r}
#2
f6_2 <- function(x)
  substitute(x+y*2)
f6_2(x=a)
```


```{r}
#3
f6_3 <- function(y) {
  x <- 5 
  substitute(x+y*2)
}
f6_3(y=4)
#compare with # 1
```


