---
title: "Chapter_15.Rmd"
output: html_document
date: "2023-04-20"
---

# S4

```{r}
library(methods) # not necessary in interactive scripts, but good practice
library(lubridate)
```

## 15.2 Basics

Use `setClass` and `new`:

```{r}
setClass("Person",
         slots = c(
           name = "character", # why in quotes?
           age = "numeric"
         ))

```

```{r}
john <- new("Person", name = "John Davis", age = NA_real_)
```


```{r}
is(john)
is(john, "Person")
```

```{r}
john@name
slot(john, "name")
```

create a setter and getter for age slot but creating generics and then methods

note that the help files say we don't need to call standardGeneric
```{r}
setGeneric("age", function(x) standardGeneric("age"))
setGeneric("age<-", function(x, value) standardGeneric("age<-"))
```

```{r}
setMethod("age", "Person", function(x) x@age)
setMethod("age<-", "Person", function(x, value) {
  x@age <- value
  x
})

age(john) <- 30
age(john)
```

### 15.2.1 Exercises

#### 1.lubridate::period() returns an S4 class. What slots does it have? What class is each slot? What accessors does it provide?

```{r}
per <- lubridate::period()
is(per)
```

```{r}
slotNames(per)
```


```{r}

tibble(name=slotNames(per),
       class=map_chr(name, ~ class(getElement(per, .x))))

```
```{r}
methods(class="Period")
```

#### 2. What other ways can you find help for a method? Read ?"?" and summarise the details.

```{r}
?"?"
```

### 15.3.6 Exercises

#### 1. Extend the Person class with fields to match utils::person(). Think about what slots you will need, what class each slot should have, and what you’ll need to check in your validity method.

```{r, eval=FALSE}
setClass("Person", 
  slots = c(
    given = "character", 
    family = "character",
#    middle = "character",
    email = "charater",
    role = "character",
    comment = "character",
#    first = "character",
#    last = "character",
    age = "numeric"
  ), 
  prototype = list(
    given = NA_character_,
    family = NA_character_,
    email = NA_character_,
    role = NA_character_,
    comment = NA_character_,
    age = NA_real_
  )
)

Person <- function(given, family=NA, email=NA, role=NA, comment=NA, age=NA) {
  age <- as.double(age)
  
  new("Person", name = name, age = age)
}

setValidity("Person", function(object) {
  lengths <- c(object@given, 
               object@family,
               object@email,
               object@role,
               object@comment,
               object@age)
  if (length(object@name) != length(object@age)) {
    "@name and @age must be same length"
  } else {
    
    TRUE
  }
})
```


#### 2. What happens if you define a new S4 class that doesn’t have any slots? (Hint: read about virtual classes in ?setClass.)

#### 3. Imagine you were going to reimplement factors, dates, and data frames in S4. Sketch out the setClass() calls that you would use to define the classes. Think about appropriate slots and prototype.

## 15.4

### 15.4.5 Exercises

#### 1. Add age() accessors for the Person class.
```{r}
setGeneric("age", function(x) standardGeneric("age"))
setMethod("age", "Person", function(x) x@age)

setGeneric("age<-", function(x, value) standardGeneric("age<-"))
setMethod("age<-", "Person", function(x, value) {
  x@age <- value
  validObject(x)
  x
})
```

```{r}
age(john)
age(john) <- 25
age(john)
```


#### 2. In the definition of the generic, why is it necessary to repeat the name of the generic twice?

once for setting the generic and one for the dispatch method (StandardGeneric).  But I don't really understand this.

#### 3. Why does the show() method defined in Section 15.4.3 use is(object)[[1]]? (Hint: try printing the employee subclass.)

Because of class hierachy

```{r}
is(per)
```


#### 4. What happens if you define a method with different argument names to the generic?

```{r, error=TRUE}
setGeneric("age<-", function(x, value) standardGeneric("age<-"))
setMethod("age<-", "Person", function(x, a) {
  x@age <- a
  validObject(x)
  x
})
```

### 15.5.5

#### 1. Draw the method graph for f(😅, 😽).

#### 2. Draw the method graph for f(😃, 😉, 😙).

#### 3. Take the last example which shows multiple dispatch over two classes that use multiple inheritance. What happens if you define a method for all terminal classes? Why does method dispatch not save us much work here?

if we define a method for all terminal classes, then the 5 other classes all have ambiguous dispatch.


### 15.6.3 Exercises

#### 1. What would a full setOldClass() definition look like for an ordered factor (i.e. add slots and prototype the definition above)?

```{r}

setClass("factor",
  contains = "integer",
  slots = c(
    levels = "character"
  ),
  prototype = structure(
    integer(),
    levels = character()
  )
)
setOldClass("factor", S4Class = "factor")

setClass("ordered",
  contains = "factor",
  slots = c(
    levels = "character",
    ordered = "logical"
  ),
  prototype = structure(
    integer(),
    levels = character(),
    ordered = TRUE
  )
)
setOldClass("ordered", S4Class = "ordered")
```

#### 2. Define a length method for the Person class.

```{r}
isGeneric("length")
getGeneric("length")
```
```{r}
setGeneric("length")
setMethod("length", "Person", function(x) length(x@name))
length(john)
```

